Testing Traversability of Protein Tunnels with Flexible Ligands Using Motion Planning

### 2017-06-30 paper.tex
conclusion

\section{Discussion \& Implementation Details}

Sampling-based motion planning is a powerful tool for computing trajectories for flexible ligands.
The proposed extensions to the RRT planner have been motivated by the necessity to cope with the narrow passage
and the high dimensionality of the flexible ligands.
By scaling-down the geometry of ligand, which is also used in other related tools~\cite{cortes2005path}, 
the relative volume of the free region of the configuration space increases, which increases the probability of placing
random samples $\qrand$ into the free region.
To further cope with the narrow passage problem, the translation part of the random samples is generated using
the 3D centerline of the tunnel.
The sampling along the tunnel centerline (in the distance defined by $\rv$) does not only lead to better growth of the tree in the tunnel, but it also limits the maximal deflection of the trajectory from the tunnel.
Therefore, the bounding box of $\C$ can be defined around the tunnel in the distance $\rv$, which is typically
smaller than the bounding box of the whole protein.
This, together with the scaled-down ligand, increases the relative volume of the free configuration space and helps the RRT to
reach the goal configuration.

The runtime of the planner is currently given mainly by the collision-detection.
As the ligand is assumed to travel only along the tunnel centerline (in the distance defined by $\rv$),
only atoms located in the vicinity of the tunnel can be considered for collision detection.
Fast collision-detection can be achieved either using fast hierarchical data structures as OBB (implemented, e.g., in~\cite{ozcollide})
    or using collision-detection system developed directly for proteins~\cite{ruiz2005biocd}.

We assume that a conformation $l \in \L$ can be switched to all other conformations in $\L$.

>>>>

\section{Discussion}

Trajectories computed for a ligand moving in a tunnel can reveal such properties that cannot be computed
only based on the geometry of the tunnels.
For example, possible detours from the tunnel can be identified, which suggest that even a tunnel with a small
bottleneck can be traversed.
The computed trajectories however cannot be considered as `real', as real ligands move inside proteins under influnce
of many forces.

This work assumes that the tunnels are computed on a static protein with the ligands, which brings several limitations.
First of all, proteins are dynamic structures and consequently also the tunnels are dynamic: their move, merge or even disappear due
to motions of protein atoms.
Currentt works in protein engineering utilizing tunnels usually employ static tunnels only, or 
various tunnels are detected separately in each frame of molecular dynamics and the tunnels are aggregated to a single one.
One of the reason is the difficulty of computing dynamic tunnels with both grid-based and VD-based methods.

The second limitation is caused by the utilization of narrow tunnels.
The protein atoms tend to fill the internal void space therefore the tunnels identified inside proteins without ligand tend
to be very narrow.
To enable at least some motion of the ligands inside such narrow tunnels, the scaling-down of the ligand atoms is necessary and it is used also in other related tool, e.g.~\cite{cortes2005path}.
Trajectories computed for scaled-down ligands cannot be considered as real trajectories by still they show easy and difficult parts of
the tunnels.
Real ligands influence protein atoms and vice versa, which enables the ligand to pass even spaces that seem to be narrow.

Due to the above reasons, the computed trajectories can be considered as too optimistic
(e.g. because they are computed on such a tunnel that is wide, but will be closed due to molecular dynamics),
or too pesimistic (e.g. the trajectory is not found because of narrow passage around the tunnel, which can in fact be
opened in molecular dynamics).

Despite these limitations, testing the tunnel traversability using motion planning technique can provide chemists more information
than the simple bottleneck radius, which is used nowadays.
To better utilize the results of motion planning, it is necessary to observe behavior of ligands in many simulations with various
proteins.

To the best knowledge of the authors, this is the first work attempt to analyze traversability of ligands in protein tunnels.
Most related works, e.g. those based on motion planning, are usually focused on protein folding, or finding arbitrary pathways
from proteins~\cite{cortes2010simulating,guieysse2008structure}.
We approached

The runtime of the planner is currently given mainly by the collision-detection.
As the ligand is assumed to travel only along the tunnel centerline (in the distance defined by $\rv$),
only atoms located in the vicinity of the tunnel can be considered for collision detection.
Fast collision-detection can be achieved either using fast hierarchical data structures as OBB (implemented, e.g., in~\cite{ozcollide})
    or using collision-detection system developed directly for proteins~\cite{ruiz2005biocd}.
We assume that a conformation $l \in \L$ can be switched to all other conformations in $\L$.
